Toward theories of deliberated preferences

### 2020-11-12 simple.tex
Mostly: hypo of memory exemplified and discussed

	Assume that the theories disagree, thus, that $\alt_1$ is satisfying and $\alt_2$ is not (say, it fails on the speed criterion), but $v(\alt_2) > v(\alt_1)$. To determine which theory is right, it is insufficient to simply observe what an individual $i \in I$ effectively chooses. This would merely represent her unreflective behavior. The approach adopted in this article to capture \acp{DJ} is to declare that the winning theory (if any) is the one that exhibit arguments able to systematically convince individuals that their proposed choice is to be adopted.
	The satisfycing theory could argue as follows: “$\alt_1$ is better than $\alt_2$, because it has a satisfactory price (lower than $t_1$, an acceptable price) and a satisfactory speed (higher than $t_2$, thus an acceptable performance); whereas $\alt_2$ is not satisfactory on the speed criterion”. 
	The additive theory could use the following argument: “$\alt_2$ is better than $\alt_1$, because $\alt_2$ is cheaper than $\alt_1$, and this more than compensates for its weaker performance in terms of speed”.

	Determining which theory is right requires to present both arguments to $i$, then let her choose one alternative in $\alts$.
\end{example}
%The proposed approach starts with this intuition that the reaction of individuals to arguments should matter, and presents a general solution to the problem of capturing reflective judgments or choices. Contrary to the simplistic situation of \cref{ex:two}, this solution must allow to confront more than just two arguments or two theories. It should define precisely what is required for a judgment to be called deliberated, and determine formally which hypothesis are required for the prescriptions of a theory of \ac{DJ} to be valid.

@@

It might be useful to exemplify those remarks in the case of our running example before discussing them further.
\begin{example}[Discussion about completeness]
	It should be remarked that there is no logical contradiction in assuming that $i$ could, for example, deliberately judge that $P ≥ D$, but have no deliberated judgment about whether $D ≥ P$ or rather $P > D$ holds, thus, whether these two lotteries have equal worth or whether $P$ is strictly more valuable.  In such a case, Deliberation completeness does not hold.
	More pragmatically, there may be a decisive argument in $\allargs$ for $P ≥ D$, but no argument in $\allargs$ strong enough to convince $i$ to pick $P$ rather than $D$ against all counter-arguments, and no argument strong enough to convince $i$ that $P$ and $D$ are equally preferable. For a formal example where this would happen, assume $\ileadsto$ is such that $i$ is lead to choose only $P$ for some of the paths including $\ar_0$ towards their end, but to be indifferent for the other ones. Thus: $\forall \hist \in \allhist \suchthat \ar_0 \in \histend: \hist \ileadsto P ≥ D$, $\exists \hist \suchthat \ar_0 \in \histend \land \hist \ileadsto P > D$ and $\exists \hist \suchthat \ar_0 \in \histend \land \hist \ileadsto D ≥ P$.
\end{example}

@@

\section{Descriptive theories}
\label{sec:descriptive}
\subsection{Informal presentation}
\commentOCf{Je ne me suis pas encore décidé : descriptive theory or empirical theory ? Pour le moment, j’utilise les deux expressions de façon interchangeable.}
A normative theory defines when a given argument is decisive for a given proposition and a given individual: whenever every sequence of arguments having that argument towards the end of the sequence lead the individual to adhere to the proposition. This, in turn, under condition of accepting its normative validity \commentYM{mais c'est quoi cette "normative validiy" ? de quelles conditions parles-tu ?}, permits in principle to know something about the \ac{DJ} of an individual. But this gives no concrete means to know anyone’s \ac{DJ}. The set of finite sequence of arguments, thus of argumentative paths, is infinite. The definition of decisiveness could admittedly be modified to account for an a priori bound on the maximal length of sequences, but this would not solve the practical problem that this set would be too large to be explored thoroughly 
\commentYM{certain que ce terme existe ?}
\commentOCf{Non, thoroughfully n’existe pas, en effet}. 
Furthermore, only some argumentative paths can be tested for a given individual: the individual’s patience is limited. A supplementary problem might happen in some situations. Once an argumentative path has been presented to the individual, this cannot necessarily be undone: she might remember some of the arguments, and they may have a lasting influence on her perspective. Let me call this possibility the hypothesis of memory.
The converse hypothesis amounts to postulating that the individual forgets after some time and use an observation protocol that separates the tests in time sufficiently.
\commentYM{je comprends ce que tu veux dire ici car ça fait un bout de temps qu'on parle de ces question. Je doute cependant qu'un lecteur innocent comprenne un traitre mot}
The hypothesis of memory further reduces the possible paths that can be tested with a given individual.

@@

Such theories satisfy the two properties that are mandatory to achieve what has been announced, namely, to obtain theories of \ac{DJ} whose power rest on a weak normative axiom and on empirically falsifiable claims. First, the claim of the theory that it argues convincingly holds if and only if no falsification sequence exist (falsification sequences are defined formally here below). Second, granted normative adequacy, if indeed a theory argues convincingly, then it is valid, meaning that the proposition that it claims are \acp{DJ} are indeed \acp{DJ}. These two claims are laid out formally in \cref{sec:validity}.

This permits to ensure that, under Normative adequacy, Validity is an empirical property. But it does not give a concrete way of arbitrating between conflicting theories. A supplementary important result, described in \cref{sec:compare}, is that when two theories disagree about incompatible propositions, one of the theories at least must fail some given sequence of empirical tests, which the proof describes. These tests can be performed on a single individual and are robust against the hypothesis of memory.

>>>>

	Assume that the theories disagree, thus, that $\alt_1$ is satisfying and $\alt_2$ is not (say, it fails on the speed criterion), but $v(\alt_2) > v(\alt_1)$. To determine which theory is right, it is insufficient to simply observe what an individual $i \in I$ effectively chooses. This would merely represent her unreflective behavior. The approach adopted in this article to capture \acp{DJ} is to declare that the winning theory (if any) is the one that exhibit arguments able to systematically convince individuals that their proposed choice is to be adopted.
	The satisfycing theory could argue as follows: “$\alt_1$ is better than $\alt_2$, because it has a satisfactory price (lower than $t_1$, an acceptable price) and a satisfactory speed (higher than $t_2$, an acceptable performance); whereas $\alt_2$ is not satisfactory on the speed criterion”. 
	The additive theory could use the following argument: “$\alt_2$ is better than $\alt_1$, because $\alt_2$ is cheaper than $\alt_1$, and this more than compensates for its weaker performance in terms of speed”.

	The basic intuition on which this article is based is that determining which theory is right requires to present both arguments to $i$, then let her choose one alternative in $\alts$, to see which argument she finds more convincing.
\end{example}
%The proposed approach starts with this intuition that the reaction of individuals to arguments should matter, and presents a general solution to the problem of capturing reflective judgments or choices. Contrary to the simplistic situation of \cref{ex:two}, this solution must allow to confront more than just two arguments or two theories. It should define precisely what is required for a judgment to be called deliberated, and determine formally which hypothesis are required for the prescriptions of a theory of \ac{DJ} to be valid.

@@

It might be useful to exemplify those remarks in the case of our running example before discussing them further.
\begin{example}[Discussion about completeness]
\label{ex:incompl}
	It should be remarked that there is no logical contradiction in assuming that $i$ could, for example, deliberately judge that $P ≥ D$, but have no deliberated judgment about whether $D ≥ P$ or rather $P > D$ holds, thus, whether these two lotteries have equal worth or whether $P$ is strictly more valuable.  In such a case, Deliberation completeness does not hold.
	More pragmatically, there may be a decisive argument in $\allargs$ for $P ≥ D$, but no argument in $\allargs$ strong enough to convince $i$ to pick $P$ rather than $D$ against all counter-arguments, and no argument strong enough to convince $i$ that $P$ and $D$ are equally preferable. For a formal example where this would happen, assume $\ileadsto$ is such that $i$ is lead to choose only $P$ for some of the paths including $\ar_0$ towards their end, but to be indifferent for the other ones. Thus: $\forall \hist \in \allhist \suchthat \ar_0 \in \histend: \hist \ileadsto P ≥ D$, $\exists \hist \suchthat \ar_0 \in \histend \land \hist \ileadsto P > D$ and $\exists \hist \suchthat \ar_0 \in \histend \land \hist \ileadsto D ≥ P$.
\end{example}

@@

\section{Descriptive theories}
\label{sec:descriptive}
\subsection{Informal presentation}
\commentOCf{Je ne me suis pas encore décidé : descriptive theory or empirical theory ? Pour le moment, j’utilise les deux expressions de façon interchangeable.}
A normative theory defines when a given argument is decisive for a given proposition and a given individual: whenever every sequence of arguments having that argument towards the end of the sequence lead the individual to adhere to the proposition. This, in turn, under condition of accepting its normative validity \commentYM{mais c'est quoi cette "normative validiy" ? de quelles conditions parles-tu ?}, permits in principle to know something about the \ac{DJ} of an individual. But this gives no concrete means to know anyone’s \ac{DJ}. The set of finite sequence of arguments, thus of argumentative paths, is infinite. The definition of decisiveness could admittedly be modified to account for an a priori bound on the maximal length of sequences, but this would not solve the practical problem that this set would be too large to be explored thoroughly 
\commentYM{certain que ce terme existe ?}
\commentOCf{Non, thoroughfully n’existe pas, en effet}. 
Furthermore, only some argumentative paths can be tested for a given individual: the individual’s patience is limited.

A supplementary problem might happen in some situations. Once an argumentative path has been presented to the individual, this cannot necessarily be undone: she might remember some of the arguments, and they may have a lasting influence on her perspective. Let me call this possibility the hypothesis of persistence, and the converse hypothesis, the hypothesis of immediacy.
Under the hypothesis of persistence, it is impossible to test twice the same path. Tests of the stability of the opinion of $i$ towards a given set of arguments can still be performed, as the next example indicates.
The converse hypothesis amounts to postulating that the individual forgets after some time and use an observation protocol that separates the tests in time sufficiently.
\commentYM{je comprends ce que tu veux dire ici car ça fait un bout de temps qu'on parle de ces question. Je doute cependant qu'un lecteur innocent comprenne un traitre mot}
The hypothesis of memory further reduces the possible paths that can be tested with a given individual.
An illustration of some consequences of these possibilities on our running example may be useful. 
\begin{example}[Hypothesis of memory]
	 The definition of $\ileadsto$ as given so far implicitly assumes that a given path can be tested at most once with a given individual, as it does not indicate what to conclude when a path $\hist$ leads $i$ to adopt a proposition $\prop$ only some of the time. This is justified under the hypothesis of memory. 
	 Recall that in \cref{ex:incompl}, $\hist \ileadsto P ≥ D$ and $\hist' \ileadsto D > P$ for some $\hist, \hist' \in \allhist$. If we further assume that $\range(\hist) = \range(\hist')$, in other words, that these two paths use the same arguments, this illustrates the remark that instability of $i$ towards some set of argument can be observed even under the hypothesis of memory. Consider $\hist = \left<\ar_0, \ar_1\right>$ and $\alpha' = \left<\ar_0, \ar_1, \ar_0\right>$ as an example. As $\hist'$ is a sequence that starts with $\hist$, it is possible to test these two paths on $i$, even under the hypothesis of memory.

	If, on the contrary, it is considered that the individual forgets over time, we must define $\ileadsto$ differently. Define $T$ a series of time steps, and given $t \in T$, define $\hist \ileadsto^t \prop$ as the observation of the adoption of $\prop$ by $i$ when having been presented the arguments in $\hist$ at time $t$. Define then $\hist \ileadsto \prop$ iff $\forall t \in T: \hist \ileadsto^t \prop$. Under the hypothesis of immediacy, it could be that $\exists \hist \suchthat \hist \nileadsto P ≥ D \land \hist \nileadsto D > P$.
\end{example}

@@

Such theories achieve what has been announced, namely, their conclusions rest solely on a weak normative axiom and on empirically falsifiable claims. 
Granted normative adequacy, if no falsification sequence exist for a given theory of \ac{DJ}, then it is valid, meaning that the propositions that it claims are \acp{DJ} are indeed \acp{DJ}.
\Cref{sec:validity} defines formally the notion of falsification sequence and proves this central claim.
The practical meaningfulness of this result is not affected by the hypothesis of memory, under condition that the set $I$ be large enough: we can always then perform falsification tests on different individuals for each path that has to be tested, so that memory of individuals does not matter. Exploring settings where $I$ is too small to apply this strategy is left aside in this article.

This result permits to ensure that, under Normative adequacy, Validity is an empirical property. But it does not give a concrete way of arbitrating between conflicting theories. A supplementary important result, described in \cref{sec:compare}, is that when two theories disagree about incompatible propositions, one of the theories at least must fail some given sequence of empirical tests, which the proof describes. These tests can all be performed on a single individual and are robust against the hypothesis of memory, which suggest that arbitration between theories could be performed reasonably easily.
